A meta-analysis of fungicide effects on mungbean powdery mildew

### 2021-12-28 03_Yield_MetaAnalysis.Rmd
Clean linebreaks, <br>, reformat code, etc.

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Grain yield meta-analysis

@@

### Define Trial
```{r Yield_define_trial}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(trial = paste(trial_ref,
                        year,
                        location,
                        host_genotype,
                        row_spacing,
                        sep = "_")) 
```

@@

`early` and `late` categories should be divided by a number close to the median days after sowing when signs of the disease were first observed to ensure roughly even numbers in each category.
```{r median onset}
PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   pull(first_sign_days) %>%
   median()
```

We will choose 42 days as this is the average days to flowering in mungbean. 
In addition, Kelly et. al. [-@Kelly2017a] noted that yield losses are higher in cultivar Berken when disease establishes before flowering.

@@

```{r}
anova(PM_mv, PM_mv_dp)
```
This comparison tells us the meta-analysis including `d_pressure` explained more residual heterogeneity, as shown by the lower QE value, and has a better fit, indicated by the higher log-likelihood with the additional `d_pressure` parameter.
However, the lower AIC in the 'Full' model indicates that retaining the model including the additional `d_pressure` variables could be a reasonable decision.
In addition the chi-squared test indicates that the `d_pressure` model is better as indicated by the `pval` 0.0468.

@@

To check the stability in the network analysis we will rerun the chosen model 15 times. 
Each run, one of the 15 trials will be excluded from the analysis to detect if any estimates varied significantly indicating an unstable network meta-analysis.
```{r stability, cache=TRUE}
stab_PM_mv <- 
   lapply(unique(PM_dat_Y$trial_ref), function(Trial){

      #subset to remove a trial
      dat <- filter(PM_dat_Y, trial_ref != Trial)

      PM_mv <- rma.mv(
         yi = grain_yield.t.ha.,
         V = vi,
         mods = ~ spray_management,
         method = "ML",
         random = list(~ spray_management | trial, ~ 1 | id),
         struct = "UN",
         control = list(optimizer = "optim"),
         data = dat
      )
   })

length(stab_PM_mv)
est <- lapply(stab_PM_mv, function(e1){
   data.frame(treat = rownames(e1$b),
              estimates = e1$b,
              p = as.numeric(reportP(e1$pval, P_prefix = FALSE,AsNumeric = TRUE)))
   })

@@

stab_df %>%
   ggplot(aes(Tnum, estimates, colour = treat))+
   geom_line(size = 1)

stab_df %>%
   ggplot(aes(Tnum, p, colour = treat))+
   geom_line(size = 1)+
   scale_y_log10()

stab_df %>%
   group_by(treat) %>%
   summarise(minEst = min(estimates),
             medEst = median(estimates),
             maxEst = max(estimates),
             minP = min(p),
             medP = median(p),
             maxP= max(p))
stab_df[Tnum == 9]
```

Trial 9 at missen flats changes the estimates the most when excluded due to the high number of reps
```{r}
unique(PM_dat_Y$trial_ref)[9]
```

### Between study heterogeneity
Calculation of I^2
```{r}
W <- diag(1/(PM_dat_Y$vi))
X <- model.matrix(PM_mv)
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
100 * sum(PM_mv$sigma2) / (sum(PM_mv$sigma2) + (PM_mv$k-PM_mv$p)/sum(diag(P)))

100 * PM_mv$sigma2 / (sum(PM_mv$sigma2) + (PM_mv$k-PM_mv$p)/sum(diag(P)))
```
This indicates that approximate amount of total variance attributed to trials/ clusters is 99%, which is very high. 
However given that many of the means included in the model contained an accompanying sample variance of zero, this might have been expected.

@@

```{r metafor_plot_yield}
results_mv %>%
   filter(Treatment != "control") %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c(
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, mu)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      linetype = 3
   ) +
   geom_point(aes(size = 1 / se), shape = 15) +
   geom_linerange(aes(ymin = `CI_{L}`, ymax = `CI_{U}`)) +
   coord_flip() +
   labs(caption = "Bars indicate 95% confidence intervals")+
   ylab(expression(paste(
      "Mean yield difference to control (t ha" ^ -1, ")", sep = ""
   ))) +
   scale_x_discrete(
      "Moderator variable",
      labels = c(
         expression("Late"["plus"]),
         expression("Late"["single"]),
         expression("Recommended"["plus"]),
         expression("Recommended"["single"]),
         expression("Early"["single"])
      )
   )
ggsave("paper/figures/Fig2_means_difference.png",
       height = 3,
       dpi = 500)
```

@@

```{r PM_Mv_Contrast_yield}
source("R/simple_summary.R") #function to provide a table that includes the treatment names in the contrasts
summary(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), test = adjusted("none"))
contrast_Ssum <-
   simple_summary(summary(glht(PM_mv, linfct = cbind(
      contrMat(rep(1, 6), type = "Tukey")
   )), test = adjusted("none")))
contrast_Ssum %>%
   flextable()%>%
   autofit()

```

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```
<br>  
<br>  
<br>

>>>>

---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Grain yield meta-analysis

@@

### Define Trial

```{r Yield_define_trial}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(trial = paste(trial_ref,
                        year,
                        location,
                        host_genotype,
                        row_spacing,
                        sep = "_")) 
```

@@

`early` and `late` categories should be divided by a number close to the median days after sowing when signs of the disease were first observed to ensure roughly even numbers in each category.

```{r median onset}
PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   pull(first_sign_days) %>%
   median()
```

We will choose 42 days as this is the average days to flowering in mungbean. 
In addition, Kelly et al. [-@Kelly2017a] noted that yield losses are higher in cultivar Berken when disease establishes before flowering.

@@

```{r}
mod <- anova(PM_mv, PM_mv_dp)
mod
```

This comparison tells us the meta-analysis including `d_pressure` explained more residual heterogeneity, as shown by the lower QE value, and has a better fit, indicated by the higher log-likelihood with the additional `d_pressure` parameter.
However, the lower AIC in the 'Full' model indicates that retaining the model including the additional `d_pressure` variables could be a reasonable decision.
In addition the chi-squared test indicates that the `d_pressure` model is better as indicated by the `pval` $`r mod$pval`$.

@@

To check the stability in the network analysis we will rerun the chosen model 15 times. 
Each run, one of the 15 trials will be excluded from the analysis to detect if any estimates varied significantly indicating an unstable network meta-analysis.

```{r stability, cache=TRUE}
stab_PM_mv <-
   lapply(unique(PM_dat_Y$trial_ref), function(Trial) {
      #subset to remove a trial
      dat <- filter(PM_dat_Y, trial_ref != Trial)

      PM_mv <- rma.mv(
         yi = grain_yield.t.ha.,
         V = vi,
         mods = ~ spray_management,
         method = "ML",
         random = list( ~ spray_management | trial, ~ 1 | id),
         struct = "UN",
         control = list(optimizer = "optim"),
         data = dat
      )
   })

length(stab_PM_mv)
est <- lapply(stab_PM_mv, function(e1) {
   data.frame(
      treat = rownames(e1$b),
      estimates = e1$b,
      p = as.numeric(reportP(
         e1$pval, P_prefix = FALSE, AsNumeric = TRUE
      ))
   )
})

@@

stab_df %>%
   ggplot(aes(Tnum, estimates, colour = treat)) +
   geom_line(size = 1)

stab_df %>%
   ggplot(aes(Tnum, p, colour = treat)) +
   geom_line(size = 1) +
   scale_y_log10()

stab_df %>%
   group_by(treat) %>%
   summarise(
      minEst = min(estimates),
      medEst = median(estimates),
      maxEst = max(estimates),
      minP = min(p),
      medP = median(p),
      maxP = max(p)
   )
stab_df[Tnum == 9]
```

Trial 9 at missen flats changes the estimates the most when excluded due to the high number of reps

```{r}
unique(PM_dat_Y$trial_ref)[9]
```

### Between study heterogeneity

Calculation of I^2

```{r}
W <- diag(1/(PM_dat_Y$vi))
X <- model.matrix(PM_mv)
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
100 * sum(PM_mv$sigma2) / (sum(PM_mv$sigma2) + (PM_mv$k-PM_mv$p)/sum(diag(P)))

100 * PM_mv$sigma2 / (sum(PM_mv$sigma2) + (PM_mv$k-PM_mv$p)/sum(diag(P)))
```

This indicates that approximate amount of total variance attributed to trials/ clusters is 99%, which is very high. 
However given that many of the means included in the model contained an accompanying sample variance of zero, this might have been expected.

@@

```{r metafor_plot_yield}
results_mv %>%
   filter(Treatment != "control") %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c(
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, mu)) +
   geom_hline(yintercept = seq(-0.05, 0.3, by = 0.05),
              linetype = 3) +
   geom_point(aes(size = 1 / se), shape = 15) +
   geom_linerange(aes(ymin = `CI_{L}`, ymax = `CI_{U}`)) +
   coord_flip() +
   labs(caption = "Bars indicate 95% confidence intervals") +
   ylab(expression(paste(
      "Mean yield difference to control (t ha" ^ -1, ")", sep = ""
   ))) +
   scale_x_discrete(
      "Moderator variable",
      labels = c(
         expression("Late"["plus"]),
         expression("Late"["single"]),
         expression("Recommended"["plus"]),
         expression("Recommended"["single"]),
         expression("Early"["single"])
      )
   )
ggsave("paper/figures/Fig2_means_difference.png",
       height = 3,
       dpi = 500)
```

@@

```{r PM_Mv_Contrast_yield}
source("R/simple_summary.R") #function to provide a table that includes the treatment names in the contrasts
summary(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), test = adjusted("none"))
contrast_Ssum <-
   simple_summary(summary(glht(PM_mv, linfct = cbind(
      contrMat(rep(1, 6), type = "Tukey")
   )), test = adjusted("none")))
contrast_Ssum %>%
   flextable() %>%
   autofit()
```

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```
